Efficiency

### 2018-09-24 Second.Rnw
Final Draft

% \usepackage{draftwatermark}
% \SetWatermarkText{DRAFT}
% \SetWatermarkScale{1.5}
% \SetWatermarkLightness{0.95}

@@

@@

@@

The open-water efficiency, $\eta_O$ was dwelt on substantially in my earlier report because this efficiency has the greatest degree of possible variance, (from zero, to around \%75 in some cases) and this variance occurs directly and necessarily with speed.  The remaining two coefficients describe the interaction between hull and the propulsor.  $\eta_R$ is the called the relative rotative efficiency, and describes the degree to which the rotating propulsor is more or less efficient operating in a wake, where the water is has an inconsistent velocity across the disc of the rotor.  $\eta_H$ describes net effect that the propulsor has on the hull by drawing water around its aft end at different speeds and altering the pressure forces on the hull, accounting for the potential for the propulsor to regain some of the energy inerent in the wake behind the hull.  Both of these effects can be shown to be most significanly driven by the hull shape and the arrangement of the propeller with respect to the hull (its relative diameter, and proximity), and be relatively constant with speed. As will be shown in this section, they could plausibly only vary a few percent each, which would have no impact in overturning the collapsing open-water efficiency of a pumpjet operating so far outside of design-conditions, such as a pumpjet at very low speed.

@@

The argument that the efficiency of pumpjets cannot be  assessed using their open-water efficiency is also blatantly false.  The 'open-water' condition is can be usefully thought of as the normal or 'ideal' efficiency of the propulsor, when the stream of incoming water is undisturbed, and consequently is still a very functional maximum efficiency of a system who's purpose is to smooth out the inconsistency in the flow.  In the absence of an advantageous pre-swirl, no higher efficiency could be reasonably expected to be achieved through such smoothing than would be acheived by beginning with a smooth stream in the first place.  Consequently, one can have considerable confidence that the extent to which this 'smoothing' effect might assist a pumpjet, relative to a propeller, is bounded by the extent to which it actually impedes a propellers performance.  As is amply demonstrated by abundant sources, this is a very small extent in all cases, and may actually be to the slight advantage of a propeller in some cases where a desirable pre-swirl already exists.

@@

Hull efficiency, like relative rotative efficiency, is not a true efficiency in that it is a ratio of two effects which can work in different directions, and consequently can be greater than one.  The two effects are the impact of the wake fraction on the propulsor, as well as the factor known as the 'thrust deduction' caused by the impact of the propulsor on the hull.

@@

The way that the wake-fraction and thrust deduction tend to scale together as a product of differing hull geometries in the case of submarines is amply demonstrated in the paper by \underline{Fundamentals of submarine concept design} by CAPT Harry A. Jackson \parencite{jackson1992}.  A third plot, reproduced here in Figure \ref{fig:hulleff.png} demonstrates how these two values combine to produce the hull efficiency.  As can be clearly seen, the size of the propeler relative to the hull also has an impact, and alters the hull efficiency by approximately 15\% when the propulsor size is halved or doubled.\footnote{The propulor size also impacts open water efficiency, so an global optimum may be constrained away from the hull-efficiency optimum.}  Given that one of the key features of the pumpjet is that they are smaller in total diameter to a comparable propeller, (and hence accelerate a smaller water-column to greater speeds) this provides a useful indication of the likley degree of variance that might be expected.  However, the size differences are unlikely to be nearly as large as a factor of two.  Reference \parencite{henderson1964}, titled \underline{A Method for the Design of Pumpjets} indicates on page that 15-20\% less diameter than an equivalent propeller is a suitable starting point for the optimisation of the mass-flow rate (which will be crucial for determining open-water efficiency, or its analogue).  As such, it would be reasonable to conclude that something of the order of 3\% is more likely to be the plausible degree of variation in the hull efficiency that might occur due to the selection of a pumpjet over a propeller.

@@

This equivalence leads to the common misunderstanding that accelerating ducted propellers are well suited to low-speed operations, such as those undertaken by tugs and trawlers.  This is not the case.  In reality, accelerating ducts are suited to propellers that are highly loaded (low $J$), which is the same end of the spectrum that a any propulsion system moves if it attempt to push a given vessel to higher speeds.  In reality, tugs and trawlers operate at unusually high $J$ values because their propeller size is constrained.  If a tug were to chose an ideally sized propeller to push around a 200,000 tonne bulk carrier, as it's routine job might be, it would naturally choose a similar sized propeller to that used by the 200,000 tonne ship it was attempting to push.  However, such a propeller would probably sink the tug, and if not the torque require to rotate it would likely cause it to capsize. With a constrained propeller size, a tug under load actually experience exactly the same propeller conditions that an over-powered speedboat, attempting speeds higher than its propeller was designed for, would experience.  In general, vessels which have high speed as their primary design consideration do not have such a propeller size constraint, and can choose to have an optimal propeller design which would be `lightly' loaded in comparison to a tugs, even when travelling at high speeds.  Of course, at lower speeds, the propeller would still be far less loaded again.

@@

This distinction is important in understanding the limitations of a pumpjet's claim to make the incoming stream of water less inconsistent, or more regular.  The duct intake and any stator blades in advance of the rotor can be effective in making the flow more homogenous as assessed on a macroscopic scale, but in doing so they simultaneously make the flow more turbulent at a smaller scale, since the necessarily involve passing the water over more surfaces, rather than less. This can be advantageous to the extent that very large-scale and intense vortices might be broken down into smaller-scale vortices, which appears more homogenous when assessed on the length-scale relevant to the blades of the rotors.  This could indeed have the effect of reducing the likelihood of very large pressure fluctuations impacting the blades, reducing the probability of cavitation occurring or invoking some kind of resonant beats within the propulsor.  \footnote{It is this effect that is likely referred to in comments made by Professor Joubert in a paper commissioned by DSTO in 2004 \parencite{joubert2004}.  However, with competently designed appendages and fairings, the kind of unsteady vortex shedding which would see such large-scale eddies shed into the propulsor should be safely avoided.  Substantial effort has been devoted to developing the smoothest possible wake-field for submarines for this purpose.\parencite{toxopeus2014}}

@@

@@

>>>>

\usepackage{draftwatermark}
\SetWatermarkText{DRAFT}
\SetWatermarkScale{1.5}
\SetWatermarkLightness{0.95}

@@

\section{Executive Summary}
In January 2018, I produced a detailed report\parencite{morrison2018}, commissioned by Gary Johnston from Submarines for Australia, in order to examine the question of the efficiency of a pumpjet.   Pumpjets, whilst common on nuclear submarines which can operate at high speeds (above 25kt) for sustained periods, are generally not used on diesel-electric submarine, which need to operate at very low speeds (around 10kt during transit, and under 5kt during combat operations).  Considerable emphasis had been placed on the significancw of the pumpjet proposed in successful French bid made by DCNS (now Naval Group) for Australia's Future Submarine.  I concluded that an inevitable characateristic of pumpjets - directly necessitated by the characteristics which gave them good acoustic performance at high speeds - was very low propulsive performance (efficiency) at low speeds.  This would mean that a diesel-electric submarine would suffer poorer range and endurance, and significantly worse dived range and endurance, as well as a worse indiscretion ratio due to the use of pumpjets.  Given the profound importance of these characteristics to all diesel-electric submarines, aswell as elevated importance of range and endurance to the Australian requirement, it would appear that a pumpjet can be safely regarded as entirely unsuitable for Australia's Future Submarine.

On June 7 2018, I appeared as a witness before the Senate Economic References Committee \parencite{SEN2018morrison}, where I defended and expanded upon my results, including in the presence of senior officials from Defence and Naval Group, who steadfastly refused to comment.  The Committee directed Defence to provide a written response to my paper prior to the conclusion of the Inquiry Into the Future Of Naval Shipbuilding Industy in Australia on 27th of June, which they finally produced in mid September.  Submarines for Australia had already commissioned me to write a subsequent report in reply to the responses that had appeared in responses to Questions on Notice in Parliament, and also in the media.  With Defence's submission finally received, this paper is in a position to comprehensively assess the totalality of the arguments put by Defence against the conclusions of my previous paper.

With the receipt of the written report it has become clear that Defence has no credible basis for rejecting the findings of my research, and instead has disgraced their technical credibility with utterly ridiculous misrepresentations of the relevant physics.  Their claim that pumpjets can be efficient across the entire speed range are instead based on the assumption that the resistance faced by fully submerged bodies is ``essentially constant over speed"\parencite{defence2018} \parencite{kerr2018}. This is simply untrue.  It is a bizarre, irrational claim with no basis whatsoever in physics.  It is frankly bewildering that such a claim could be made, given how easily it's falsehood can be established by even moderate research, or simple logic.  If it were true, then submarines and torpedoes would be able to elevate their speed without increased resistance, and consequently reach any speed they might choose without requiring larger motors, or more fuel, and hence attain virtually unlimited range.

The truth is that the resistance of submerged bodies in the real world scales with velocity squared.  This exponential increase in resistance explains why conventional submarines slow to extremely low speeds in order to conserve energy when dived, and has been used by other respected physicists discussing this topic \parencite{davies2017} as well as essentially all scientific literature involved hydrodynamics.  It also explains why a propulsor, which must faces one hundred times as much resistance when travelling at 30kt as it does at 3kt, behaves very differently in these regimes, and consequently has a different efficiency.  For Defence, assisted by DSTG, to claim that the relationshp between drag resistance and speed under-water is constant is an extremely shocking revelation.  In the engineering/technical community, it would be just as strange to hear that a dentist with teenage children believes in the tooth-fairy.  Not only is is their job to know otherwise, but abundant life experience would have continually proven otherwise.

An unavoidable conclusion is that claims made by Defence to Parliament in response to Questions on Notice claiming that the pumpjet can be efficient across the entire speed range are entirely baseless and extremely misleading.

Such a revelation necessitates a further question regarding whether such misleading claims where made out of complete ignorance and incompetence, or whether they were deliberately intended to deceive.  As can be seen from the chronology of public comments which this report details, it is difficult to see how Defence could hold such a naive and embarrassing view after having their attention drawn to the opposing evidence multiple times, and devoted so much resources to understanding the relevant technologies.  In 2012 Senator Johnston challenged RADM Moffit directly as to why research money was even being directed towards pumpjets, given how obviously inappropriate they were for our use, but RADM Moffit insisted they would still do the work required. In 2017 Defence revealed that a 99 page report was commissioned on pumpjets as part of \$214 million in 'de-risking' research, and that over a million dollars had subsequently been spent on pumpjet research specifically.

Notwithstanding the improbability that the incompetence of Defence is so completely pervasive that all of this research failed to yeild even a basic understanding of hydrodynamics, the probability that the necessary degree of ignorance has occurred through arrangement rather than accident can be most firmly established by three simple propositions, which are made obvious through a reading of this report:

\begin{enumerate}
\item  \textbf{The French know the answer.}  As one of only two countries in the world who have designed and produced both pumpjets and propellers, nuclear and conventional submarines, and fitted and operated each type of propulsor on both types of submarines, there can be no doubt whatsoever that Naval Group and the French Government do have well-formed opinion about the performance of pumpjets on diesel-electric submarines.  They would have extensive operational performance data from their in-service SSBNS which use the pumpjet.  They would have extremely detailed modelling, and tank-test results from the pumpjet from the almost-complete Barracuda class, which has an identical hull diameter to the proposed Shortfin Barracuda. The would have extensive operational performance data from their own Rubis Class nuclear submarines, which still use propellers, as well as the results of earlier tests, confirmed to have occurred by Naval Group Officials, where a pumpjet was fitted to a diesel-electric Agosta Class submarine.  That they have adequate performance data from all the necessary reference points to assess the likely performance of the pumpjet on our own is entirely beyond dispute.

\item  \textbf{The French gave the answer to the Australians.} It would naturally be a matter of profound neglect on the part of Australian officials to fail to ascertain the necessary performance data - both acoustic and propulsive efficiency - from the French during a Competitive Evaluation Process.  Any doubt that this occurred, however, can be reliably dismissed on account of public statements made by DCNS during the CEP claiming proudly that ``The stealth and hydrodynamic performances of pump jet propulsion are of course classified and in Australia known only to DCNS and the Australian Government."\parencite{costello2016}. Given RADM Sammut's repeated statements \parencite{SEN2016sammut} \parencite{SEN2017hanson} to Senate Estimates that  ``the way in which technical data would be transferred to Australia and the way in which intellectual property would be managed" was one of the reasons that DCNS was selected as design partner, and DCNS's clear pride in  ``the offer from France to transfer to Australia sovereign control and use of pump-jet propulsion technology" there can also be no doubt that after the conclusion of the CEP and the commencement of design, absolutely nothing would have prevented Australia seeking further performance details or reference points, or even assistance in interpretation in the case that the information provided in the CEP was found to be insufficient.

\item \textbf{The Australians can't produce the answer, and are instead denying one exists.} Instead of countering my report by claiming that they have examined test results provided by the French, (or from their own research and modelling) which demonstrates that pumpjet efficiency does not decline at low speeds as I have concluded, Defence is now disgracing themselves with a series of entirely implausible arguments which amount to arguing that no adequate reference point really exists, and therefore the answer cannot be known.  In simple terms, those arguments are:

\begin{itemize}

\item \textbf{The hull and the pumpjet need to be `tuned' to each other.} This is the longest-running argument, which has been made repeatedly by RADM Sammut in sequential years \parencite{sammut2017}, \parencite{bergmann2018}, in answers to Questions on Notice \ref{QON1}, and by Chief of Maritime at DSTG \parencite{kerr2018} as well as the final response to my paper.\parencite{defence2018}.  It constitutes an attempt to deny that other nuclear submarines, including the nuclear Barracuda, serve as an adequate reference point to assess the probable performance of the pumpjet on the Shortfin Barracuda.  It requires an enormous contortion of logic order to argue that the performance of the pumpjet we will have can't be known on account of the hull.  It would be equally fair to argue that one cannot know how a pair of shoes would perform in a given circumstance without simultaneously tuning them to a pair of socks. Whilst it is true that the two have to work in close conjunction with each other and a particularly poor match might impair performance, it is necessary (and sufficient) during concept design to assume that a competent (or even ideal) match is made in order to assess the overall suitability of the system.  The implausibility of this argument is further cemented by just how much emphasis has been given to the fact that the Shortfin Barracuda will have \textit{exactly the same} diameter hull as the nuclear cousin from which it is derived \ref{SEN2017hanson}, \parencite{kerr2017}, which ought to make adjustments to the pumpjet on account of the hull entirely unwarranted.

\item \textbf{The weight of the pumpjet makes it inappropriate for smaller diesel-electric submarines, but not for larger ones.} This argument has only occurred recently, first made by former CEO of Naval Group Australia Brent Clark \parencite{clark2018}, (to which I responded\parencite{morrison2018b}), and then later explained in more detail by Chief of Maritime at DSTG Dr David Kershaw\parencite{kerr2018}.  It is an attempt to deny that any other non-nuclear submarine, including every operating class in existence (none of which pumpjets) constitute an adequate reference point for assessing the likely performance of the pumpjet on our future non-nuclear submarine.  It is demonstrably ridiculous, since if a smaller submarine could not maintain proper balance and buoyancy with a heavy pumpjet attached to its tail, then the the submarine would be doomed to either tail-stand or sink, which would amount to a catastrophic failure if ever such a submarine were to detach from its moorings. That this is not the case is amply demonstrated by the fact that both the French and Russians have succeeded in fitting pumpjets to small submarines, and such catastrophes were avoided.  The Russian test boat, B-871 `Alrosa' still operates regularly in the Black Sea.\footnote{The Black Sea is an inland sea about half the size of NSW, which shares borders only with Romania, Bulgaria, Georgia, Ukraine and Turkey. Russia operates a fleet of seven Kilo Class Submarines in the Black Sea, mostly out of Sevastopol, the port which it aquired control of through the 2014 annexation of Crimea.  The Black Sea might thus be described appropriately as a `Russian Lake', where the endurance and range penalties imposed on Alrosa by the pumpjet might be tolerated.  None of the Kilo Class Submarines produced since Alrosa have a pumpjet, and none of the 15 Russian Kilo Class submarines operating in the Northern, Pacific, or Baltic Fleets have pumpjets.}

\item \textbf{The resistance of submerged bodies is constant with speed, unlike that of ships who encounter wave-making resistance.}  This is the most recent, and the most utterly and demonstrably false claim, made both by Dr David Kershaw of DSTG in an interview with Australian Defence Magazine \parencite{kerr2018}, and in Defence's considered written submission to the Senate Economic References Committee \parencite{defence2018}.  It constitutes a clearly desperate and severely embarrassing attempt to argue the no propulsion system of any sort as tested or demonstrated on a surface vessel serves as any kind of reference point for its performance on a submarine.  It is true that ships have the \textit{additional} resistance from wave-making resistance, but for most displacement vessels operating at design speed (and certainly all submarines) pressure drag and skin friction on the submerged part of the hull tend to dominate, both of which scale with velocity squared.  Consequently, the `advance ratio' (how far forward through the water a propeller moves for a given degree of rotation) changes enormously with speed, for a submarine just as it does for a surface ship, and efficiency consequently changes along with this.

\end{itemize}

\end{enumerate}

As noted by Andrew Davies when he contemplated the possibility that `Naval Group is pulling the wool over the eyes of a clueless Commonwealth' with the pump jet, extraordinary claims require extraordinary evidence \parencite{davies2017}.  There can be no doubt that such extraordinary evidence of cluelessness has now been supplied by the Department of Defence themselves. But this evidence, revealing impression of cluelessness so complete, is difficult to reconcile with the adamant insistence of competence and confidence in the choice, given by multiple senior officials in the face of sustained questions backed up by substantial evidence. It is so extraordinary that it demands the question of whether the appointed officials who represent the Commonwealth are indeed `clueless' or whether officials in the Navy and Department of Defence are actually conspiring with Naval Group and the French Government to pull the wool over the eyes of a clueless Parliament, as well as the properly elected members of the executive who ultimately signed off on the purchase, and the Australian public at large.

The gravity of such a conclusion is difficult to understate, given the likely nature of the motives which would attend to such a decision.  One would presume that one would not advance into the public arena a submarine design which was known to be unviable, and lead to deep embarassment and ridicule if subjected to thorough scrutiny, unless the real motive for the design or partner choice was even less capable of withstanding public scrutiny.  The possibility that this is the case must be considered very seriously, given the difficulty of reconciling recent events with the counterfactual proposition:

After the initial publication of my first report, three Questions on Notice were asked by Senator Ketter, specifically referencing my report \ref{QON1}.  Whilst Defence avoiding answering all of them directly, one of them was particularly brazenly side-stepped.  Asked ``Is Defence still confident that Naval Group still hold the opinion that the pump-jet is well suited to diesel-electric submarines?" Defence merely reiterated their intent to use the pumpjets as offered, strictly avoiding mentioning anyone's opinion, either their own or Naval Group's.  If the answer actually was what we would all assume it should be, that should be a very easy answer to directly.  A simple `yes' would do just fine.  If the answer is `no', then we must assume that at the time of writing (presumeably before the return date of 3 April, though Defence isn't always timely) that Defence an inkling (at the very least) of what Naval Group undoubtedly knew all along - that pumpjets are entirely unsuitable for diesel-electric submarines.  In avoiding answering the question directly, Defence was not just misleading the parliament, but clearly acting to shelter Naval Group from subsequent accusations of fraudulent or misleading conduct during the CEP.

Giving Defence the most generous possible interpretation, one might allow the possibility that Defence were just coming to grips with being completely hood-winked during the CEP by their chosen design-partner, and perhaps they felt that their wrath at being deliberately deceived ought not be publicly vented, or at least not at this stage.  One might, at a stretch, allow that maintaining the opportunity for the partner to save-face was still a useful bargaining-chip that our Defence chiefs to use in order to bring a potentially treacherous contractor to heel, along with a range of contractual penalties, as well as the commercial leverage the Commonwealth would have by being able to re-introduce competition into the design process as many have suggested ought be done in any case \parencite{stanford2017}, or not proceed to a build contract on completion of the design.

The state of commercial negotiatons between the Commonwealth and Naval Group, as revealed in the Senate Economic References Committee \parencite{SEN2018morrison}, reveal that Defence was moving on exactly the opposite tack - shocking Senators with the revelation that Defence was actually moving instead to contract immediately with DCNS for the full construction of all 12 submarines, to the complete exclusion of the Australian Submarine Corporation, within coming weeks, effectively wiping out the opportunity for any competitive tension at any stage of the ~\$50 billion project.

In light of these events, I find the most plausible explanation to be the worst:
\begin{itemize}
\item The propulsive performance of pumpjets do indeed decline to very low efficiencies at low speeds, making them entirely unsuitable for a diesel-electric submarine which operates almost entirely at such speeds, broadly in line with the findings of my first report.
\item Some group of individuals, including senior French officials from Naval Group as well as the Australian Government, were aware of this fact at the time of the Competitive Evaluation Process, and remains so today.
\item That group has deliberatly cultivated the impression that pump-jets would confer a substantial technological edge to Australia's Future Submarine, knowing that would not be the case, either by virtue of the pump-jets not ultimately being used, or because if they were they would confer a disadvantage.
\item That impression was necessary in order to defend of a choice which selected a design partner who's proposal was both more expensive, and slower to deliver that their competitors, for which no more defensible reason could be found.
\end{itemize}

It appears that the pumpjet idea was simply a lead-character used to construct the fiction that France's vaunted heritage as a designer of nuclear submarines offers a substantial technical advantage to a non-nuclear submarine.  There is almost no chance that it will be included in the detailed design.  In reality, without a nuclear reactor, there is barely any relevant technology or experience (other than the combat system, mandated from a 3rd party) which is usefully transferrable.  The extent to which the bottom of the barrel needed to be scraped was illustrated when RADM Sammut was challenged on the benefit of the nuclear design, other than pump-jets \ref{SENhanson}, and he was forced to refer to the `diameter of the hull' as being a key learning, a prospect which would be much more easily subjected to public ridicule had it been more substantially promoted.\footnote{Amongst the many engineering challenges presented in designing a new submarine, ensuring the structural integrity of the hull at a new diameter is one of the most well-understood and predictable.  The Collins Class was evolved from a submarine of smaller hull diameter, and this aspect of the design and development encountered no problems.  In 2006, DSTO commissioned a paper suggesting that the Collins Class replacement be designed at a wider hull, which didn't correspond to that of the Barracuda.}  In contrast, those countries whos own navies relied entirely upon diesel-electric submarines would likely boast significant advantages in technologies which are genuine game-changers for diesel-electric submarines, such as advanced composite propellers, Air-Independent Propulsion, and Lithium-Ion batteries, which we are now told won't be included in the Future Submarine \textit{because} of the pumpjet\parencite{bergmann2018}.

Australians have a right to know what the true motive of Defence was in choosing the French as a design partner.  There can be no doubt that the pumpjet played no legitimate part in it.  Any claims to the contrary are made either out of obscene incompetence, as Defence's latest response might have us believe, or more likely are made with the intent of being misleading.

@@

\subsection{RADM Gregory Sammut, Head of SEA 1000 before the Senate}

\begin{tcolorbox}[width=0.9\textwidth, center, colback=red!5!white,colframe=red!75!black, title= \textbf{6 May 2016 - Senate Estimates}]
\textit{Senator Conroy asks Rear Admiral Greg Sammut (then director of SEA 1000) about the propulsion system of the proposed submarine.  Full exchange in \ref{SENsammut2}. Emphasis added.}

\textbf{Senator CONROY:}  The French do not have this propulsion system on a conventional submarine anywhere, do they?

\textbf{Rear Adm. Sammut:}  They do not have a nuclear reactor on a conventional submarine.

\textbf{Senator CONROY:}  No, the propulsion system they are proposing for the Shortfin Barracuda is not powered by diesel; it is only nuclear powered.

\textbf{Rear Adm. Sammut:}  DCNS has proposed to us a conventionally-powered submarine that includes—I think you are trying to get to the form of propeller we are talking about here, Senator.

\textbf{Senator CONROY:}  Yes.

\textbf{Rear Adm. Sammut:}  It is a pump jet, which is what DCNS has said is part of their proposal. That is just one element of the entire propulsion system. We were discussing before the fact that there are many elements to the propulsion system. There is the propeller, there is the motor that drives the propeller and the diesels that charge the battery that drives the motor. All of those factors combined lead to the propulsion system that is fitted to a submarine . When we consider all of those elements we look at what each contributes to the likely noise signature of the submarine and make an assessment about the overall performance of the submarine as far as its stealth is concerned.

\textbf{Senator CONROY:}  I have read that the signature of the nuclear-powered sub with the propulsion system propeller that we are taking is very good at high speed. But at the speeds that a conventional sub will go-which by definition are considerably less than the speeds a nuclear-powered one achieves-there is a difference in the signature. Could you comment on that?

\textbf{Rear Adm. Sammut:}  \textbf{Only to say that there is an impact across all speed ranges of the propulsion system. It does not purely pertain to high speed.}

\textbf{Senator CONROY:}  But what I have read is that the propulsion system is, if you like, designed and attuned to perform against the nuclear-powered aspect, and that they swap the powering out and keep the same propulsion system, but they have not adapted it in any specific way. I am inviting you to comment. The two parts work beautifully together, but the conventional power that we plan to use does not create as successful a signature at the lower speeds—so above 20 knots versus below 20 knots. Would you like to comment?

\textbf{Rear Adm. Sammut:}  Only to say what I said before. There is an advantage across all the speed ranges in the technologies that are being proposed by DCNS. \textbf{You are right: there has to be a degree of adaption of any propeller that is designed for a submarine , whether it is a normal propeller or whether it is a pump jet.} That always has to take place. The principles that underpin the quietness of a pump jet apply in any case.

Only weeks after the announcement of the success of the DCNS bid to be the design partner for SEA 1000, specific questions about the speed-suitability of the propulsion system are being raised by senators.  In this particular line of questioning, Senator Conroy implies that the acoustic advantages of the pumpjet may not apply as well at the speeds that a conventional submarine could operate at, which is consistent with findings in my earlier report that the acoustic advantage would essentially be constrained to circumstances where high speeds would cause cavitation on an open propeller. In his responses, RADM Sammut claims that the advantages apply at all speeds.  He also concedes that some modification of the pumpjet will be required for the purpose, in response to a question about the suitability of the speed.  He is not, however, specific about whether it is the speed requirement, or the shape of the hull, which drives this need for adjustment, though the context would seem to imply the former.  He also here clearly discusses a pumpjet as being a special type of propeller, rather than a completely unrelated system which bears no comparison.

@@

The open-water efficiency, $\eta_O$ was dwelt on substantially in my earlier report because this efficiency has the greatest degree of possible variance, (from zero, to around 75\% in some cases) and this variance occurs directly and necessarily with speed.  The remaining two coefficients describe the interaction between hull and the propulsor.  $\eta_R$ is the called the relative rotative efficiency, and describes the degree to which the rotating propulsor is more or less efficient operating in a wake, where the water is has an inconsistent velocity across the disc of the rotor.  $\eta_H$ describes net effect that the propulsor has on the hull by drawing water around its aft end at different speeds and altering the pressure forces on the hull, accounting for the potential for the propulsor to regain some of the energy inerent in the wake behind the hull.  Both of these effects can be shown to be most significanly driven by the hull shape and the arrangement of the propeller with respect to the hull (its relative diameter, and proximity), and be relatively constant with speed. As will be shown in this section, they could plausibly only vary a few percent each, which would have no impact in overturning the collapsing open-water efficiency of a pumpjet operating so far outside of design-conditions, such as a pumpjet at very low speed.

@@

The argument that the efficiency of pumpjets cannot be  assessed using their open-water efficiency is also blatantly false.  The `open-water' condition is can be usefully thought of as the normal or `ideal' efficiency of the propulsor, when the stream of incoming water is undisturbed, and consequently is still a very functional maximum efficiency of a system who's purpose is to smooth out the inconsistency in the flow.  In the absence of an advantageous pre-swirl, no higher efficiency could be reasonably expected to be achieved through such smoothing than would be acheived by beginning with a smooth stream in the first place.  Consequently, one can have considerable confidence that the extent to which this `smoothing' effect might assist a pumpjet, relative to a propeller, is bounded by the extent to which it actually impedes a propellers performance.  As is amply demonstrated by abundant sources, this is a very small extent in all cases, and may actually be to the slight advantage of a propeller in some cases where a desirable pre-swirl already exists.

@@

Hull efficiency, like relative rotative efficiency, is not a true efficiency in that it is a ratio of two effects, either of which can be greater than the other, and consequently can be greater than one.  The two effects are the impact of the wake fraction on the propulsor, as well as the factor known as the `thrust deduction' caused by the impact of the propulsor on the hull.

@@

The way that the wake-fraction and thrust deduction tend to scale together as a product of differing hull geometries in the case of submarines is amply demonstrated in the paper by \underline{Fundamentals of submarine concept design} by CAPT Harry A. Jackson \parencite{jackson1992}.  A third plot, reproduced here in Figure \ref{fig:hulleff.png} demonstrates how these two values combine to produce the hull efficiency.  As can be clearly seen, the size of the propeller relative to the hull also has an impact, and alters the hull efficiency by approximately 15\% when the propulsor size is halved or doubled.\footnote{The propulor size also impacts open water efficiency, so an global optimum may be constrained away from the hull-efficiency optimum.}  Given that one of the key features of the pumpjet is that they are smaller in total diameter to a comparable propeller, (and hence accelerate a smaller water-column to greater speeds) this provides a useful indication of the likley degree of variance that might be expected.  However, the size differences are unlikely to be nearly as large as a factor of two.  Reference \parencite{henderson1964}, titled \underline{A Method for the Design of Pumpjets} indicates on page that 15-20\% less diameter than an equivalent propeller is a suitable starting point for the optimisation of the mass-flow rate (which will be crucial for determining open-water efficiency, or its analogue).  As such, it would be reasonable to conclude that something of the order of 3\% is more likely to be the plausible degree of variation in the hull efficiency that might occur due to the selection of a pumpjet over a propeller.

@@

This equivalence leads to the common misunderstanding that accelerating ducted propellers are well suited to low-speed operations, such as those undertaken by tugs and trawlers.  This is not the case.  In reality, accelerating ducts are suited to propellers that are highly loaded (low $J$), which is the same end of the spectrum that a any propulsion system moves in if it attempts to push a given vessel to higher speeds.  In reality, tugs and trawlers operate at unusually high $J$ values because their propeller size is constrained.  If a tug were to chose an ideally sized propeller to push around a 200,000 tonne bulk carrier, as it's routine job might be, it would naturally choose a similar sized propeller to that used by the 200,000 tonne ship it was attempting to push.  However, such a propeller would probably sink the tug, and if not the torque require to rotate it would likely cause it to capsize. With a constrained propeller size, a tug under load actually experience exactly the same propeller conditions that an over-powered speedboat, attempting speeds higher than its propeller was designed for, would experience.  In general, vessels which have high speed as their primary design consideration do not have such a propeller size constraint, and can choose to have an optimal propeller design which would be `lightly' loaded in comparison to a tugs, even when travelling at high speeds.  Of course, at lower speeds, the propeller would still be far less loaded again.

@@

This distinction is important in understanding the limitations of a pumpjet's claim to make the incoming stream of water less inconsistent, or more regular.  The duct intake and any stator blades in advance of the rotor can be effective in making the flow more homogenous as assessed on a macroscopic scale, but in doing so they simultaneously make the flow more turbulent at a smaller scale, since it necessarily involve passing the water over more surfaces, rather than less. This can be advantageous to the extent that very large-scale and intense vortices might be broken down into smaller-scale vortices, which appears more homogenous when assessed on the length-scale relevant to the blades of the rotors.  This could indeed have the effect of reducing the likelihood of very large pressure fluctuations impacting the blades, reducing the probability of cavitation occurring or invoking some kind of resonant beats within the propulsor.  \footnote{It is this effect that is likely referred to in comments made by Professor Joubert in a paper commissioned by DSTO in 2004 \parencite{joubert2004}.  However, with competently designed appendages and fairings, the kind of unsteady vortex shedding from appendages which would see such large-scale eddies shed into the propulsor should be safely avoided.  Substantial effort has been devoted to developing the smoothest possible wake-field for submarines for this purpose.\parencite{toxopeus2014}}

@@

Having considered in some detail the nature of the three substantive arguments given in opposition to my findings, it is clear that none actually provide any credible evidence to oppose my conclusions.  However, a clear similarit or parallel between the three can be found.  Each of them respectively actually comprise an argument against an obvious reference point being used to assess the efficiency of the pumpjet.

The argument that resistance is constant for underwater bodies, since they encounter no wave-making resistance, seeks to oppose any use of propulsive data derived from propulsors which might be used on ships, including for highly comparable systems (waterjets), or even directly comparable systems (pumpjets, or decelerating ducted propellers).

The argument that the size of the submarine is a game-changer, attempts to dismiss the possibility that any experience with other diesel-electric submarines constitutes and adequate reference-point.

The argument that the hull and the pumpjet need to be tuned together is clearly an argument that is intended to dismiss the possibility than any other nuclear submarine, including the Barracuda from which the Shortfin Barracuda is derived, serves as an adequate reference point.

It is indeed remarkeable that none of the arguments made actually involve pointing to a better reference point than what I have been able to use from open-source literature, or announce that any results from Naval Group directly contradict the findings of my earlier report.  It is a relatively simple proposition that Naval Group, as one of only two countries in the world to have fitted both pumpjets and propellers to both diesel and nuclear submarines, has in its possession the best data possible to serve as reference-points for our submarine.  Detailed modelling, and tank testing from the pumpjet on the Barracuda would certainly provide the best reference, but experience from the in-service Rubis Class (which use propellers) as well as the in-service SSBNs would doubtless also provide adequate data.

As I have argued, I think that there actually are adquate reference points in open-source literature regarding comparable propulsion systems, designed for comparable purposes, employed in relevant and comparable, in order to gain what is really only a very broad, but clear conlcusion: that propulsive efficiency declines inevitably and dramatically at low speeds.

Taking it as given that the French do have posession of adquate of far more detailed performance data, derived from even closer reference points than what I can glean from public literature, one must ask whether that data actually confirms, or contradicts my findings.  Given public claims from DCNS during the CEP that information regarding both the hydrodynmic and acoustic performance of the pumpjet were in Australian government hands in Australia \parencite{costello2016} there can be little doubt Australian officials have access to such information.  Indeed, it would be an inexcuseable oversight if they didn't obtain that information during the CEP.

This poses a seemingly irreconcileable tension between the desparate, implausible arguments currently being made by Defence, and the possibility that they are being made because Defence genuinely doesn't know better, and genuinely believes them to be true.  It seems far more probable that Defence actually is in posession of clear, conclusive evidence which broadly supports the conclusions of my report, and is now deliberately concealing that evidence, and attempting to excuse why they have not drawn the obvious conclusions that any sensible person normally would from that evidence - that a pumpjet is entirely unsuited to a diesel-electric submarine.

@@

\subsection{Hansard Extract, Senate Estimates - 6th May 2016 \cite{SEN2016sammut}}\label{SENsammut2}

\textbf{Senator CONROY:}  The joint media release, Minister, states, 'Subject to discussions on commercial matters, the design of the Future Submarine» with DCNS will begin this year,' which I think is what Admiral Sammut is indicating. Is it fair to say that the process so far, up until the announcement, was about identifying a design partner—not a specific submarine design—and that the specifics, such as cost, schedule and performance, are still to be worked out? I am just going off the press release that you issued, Minister.

\textbf{Senator Payne:}  As you said in your previous question to Admiral Sammut, we are at the beginning of the design process, not the end. So, in selecting DCNS, the evaluation process identified that proposition, which they submitted, as the one which best met the capability requirements that Australia has for a regionally superior submarine . We will now proceed to work with the DCNS, as Admiral Sammut has outlined, to develop that design.

\textbf{Senator CONROY:}  There has been a lot of robust public debate about the decision. I want to offer Admiral Sammut the opportunity to outline to the committee, and to the broader public who pay attention to these things, the basis on which the decision was made. We have established that we do not have the finalised costs, we do not have a finalised production schedule and we do not have a mature design, so what were the factors that you based your decision around in choosing the French submarine ?
